---
title: "p8105_hw1_dz2399"
author: "DZ"
date: "September 29, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1

### Read in the data according to instuctions
```{r}
df_nyc <- read_csv(file='./NYC_Transit_Subway_Entrance_And_Exit_Data.csv') %>%
  janitor::clean_names() %>%
  select(line, starts_with('station'),starts_with('route'),entry, vending,ada,entrance_type) %>% 
  select(-station_location) %>% 
  mutate(entry = recode (entry, 'YES' = T, 'NO'=F))

dim(df_nyc)
```

 The dataset contains the following variables: line, station_name, station longitute and latitude, the route that the train serves, entry and entry types, vending, and ada. The data cleaning process is that: firstly, read in the data from the directory and conduct a column-name-cleaning with the function clean_names; then, select the desired variables; lastly, recode the entry variable with yes equals to T and no to F. The dimension of the resulting dataset is `r dim(df_nyc)`. This is not a tidy data, because the route variables are spreadout to multiple columns rather than a single column.

### How many distinct stations are there?
```{r}

```


